Theoretical aspects in gravity modeling

### 2018-04-22 manuscript/letter.tex

In a local or regional study, geophysicists commonly use a topocentric 
Cartesian coordinate system (TCS) with origin at a point $P$ on or 
close to the Earth's surface and axes $x$, $y$ and $z$ (Figure 
\ref{fig:TCS}). In the TCS, the axes $x$ and $y$ are parallel to 
the unit vectors $\hat{\mitbf{v}}_{P}$ $\hat{\mitbf{w}}_{P}$, respectively,
whereas the $z$-axis is opposite to the unit vector 
$\hat{\mitbf{u}}_{P}$ (equation \ref{eq:unit-vectors}) and points downward.

Consider a TCS (Figure \ref{fig:TCS}) with origin at a point $P$ with
coordinates $(X_{P}, Y_{P}, Z_{P})$ referred to the GCS (Figure \ref{fig:GCS-GGS}).
In this case, the relationship between the coordinates $(X, Y, Z)$ of a 
point in the GCS and the coordinates $(x, y, z)$ of the same point in the 
TCS is given by:
\begin{equation}
\begin{bmatrix}
x \\
y \\
z 
\end{bmatrix} =
\mitbf{R}^{\top} \begin{bmatrix}
X - X_{P} \\
Y - Y_{P} \\
Z - Z_{P}
\end{bmatrix} \: ,
\label{eq:GCS2TCS}
\end{equation}
where $\mitbf{R}$ is an orthogonal matrix whose columns
are defined by the unit vectors $\hat{\mitbf{u}}$, $\hat{\mitbf{v}}$ and
$\hat{\mitbf{w}}$ (equation \ref{eq:unit-vectors}) as follows:
\begin{equation}
\mitbf{R} = 
\begin{bmatrix}
\hat{\mitbf{v}} & \hat{\mitbf{w}} & -\hat{\mitbf{u}}
\end{bmatrix} \: .
\label{eq:R}
\end{equation}

@@

The gravity anomaly is the commonly used quantity in applied 
geophysics. It is defined as the difference between the 
gravity at a point $Q^{\prime}$ on the geoid and the normal gravity 
at a point $Q$ on the ellipsoid surface,
both at the same geodetic latitude and longitude (Figure \ref{fig:surfaces}).
Notice that, by definition, the gravity anomaly depends on 
longitude and latitude only and is not a function of height 
\citep{barthelmes2013}. As a consequence, it is not possible to compute,
for example, the upward of a gravity anomaly.

>>>>

In a local or regional study, geophysicists commonly use a topocentric 
Cartesian coordinate system (TCS) with origin at a point $P$ on or 
close to the Earth's surface and axes $x$, $y$ and $z$ (Figure 
\ref{fig:TCS}). In the TCS, the axes $x$ and $y$ are parallel to 
the unit vectors $\hat{\mitbf{v}}_{P}$ $\hat{\mitbf{w}}_{P}$, respectively,
whereas the $z$-axis is opposite to the unit vector 
$\hat{\mitbf{u}}_{P}$ (equation \ref{eq:unit-vectors}) and points downward.
Consider, for example, a TCS (Figure \ref{fig:TCS}) with origin at a point $P$ with
coordinates $(X_{P}, Y_{P}, Z_{P})$ referred to the GCS (Figure \ref{fig:GCS-GGS}).
In this case, the relationship between the coordinates $(X, Y, Z)$ of a 
point in the GCS and the coordinates $(x, y, z)$ of the same point in the 
TCS is given by:
\begin{equation}
\begin{bmatrix}
x \\
y \\
z 
\end{bmatrix} =
\mitbf{R}^{\top} \begin{bmatrix}
X - X_{P} \\
Y - Y_{P} \\
Z - Z_{P}
\end{bmatrix} \: ,
\label{eq:GCS2TCS}
\end{equation}
where $\mitbf{R}$ is an orthogonal matrix whose columns
are defined by the unit vectors $\hat{\mitbf{u}}$, $\hat{\mitbf{v}}$ and
$\hat{\mitbf{w}}$ (equation \ref{eq:unit-vectors}) as follows:
\begin{equation}
\mitbf{R} = 
\begin{bmatrix}
\hat{\mitbf{v}} & \hat{\mitbf{w}} & -\hat{\mitbf{u}}
\end{bmatrix} \: .
\label{eq:R}
\end{equation}

@@

The gravity anomaly is the commonly used quantity in applied 
geophysics. It is defined as the difference between the 
gravity at a point $Q^{\prime}$ on the geoid (a particular 
equipotential surface of the gravity potential) and the normal gravity 
at a point $Q$ on the ellipsoid surface,
both at the same geodetic latitude and longitude (Figure \ref{fig:surfaces}).
Notice that, by definition, the gravity anomaly depends on 
longitude and latitude only and is not a function of height 
\citep{barthelmes2013}. As a consequence, it is not possible to compute,
for example, the upward continuation of a gravity anomaly.
However, many authors in the literature have computed the upward 
continuation of gravity anomalies. All these authors have implicitly 
considered that the gravity anomaly is an approximation of 
the gravity disturbance, which in turn can be represented by a 
harmonic function that can be continued upward.
